---
title: "1. DAIE Pair Project"
author: "Joe O'Regan and Michael Galbraith"
date: "2022-12-12"
output:
  html_document:
    toc: yes
    toc_depth: '2'
    df_print: paged
  html_notebook:
    theme: darkly
    toc: yes
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load & View, echo=TRUE}

library(readr)
data5 <-read_csv("daie_ca3_data_5.csv")
View(data5)
#head(data5, 10)
```

**Repo:** <https://github.com/joeaoregan/2022_DAIE_GCA_JOR_MG>

# Title
# Table of Contents
# Abstract

summary of the contents

## Aim and Rationale

## Participants and Setting

## Experiment Design

## Results Gathering

## Major Findings

## Findings / Implications


# Introduction

## Topic and Context
MG:
Post‐traumatic stress disorder (PTSD) is arguably the most common psychiatric disorder to arise after exposure to a traumatic event. Trauma‐focused cognitive behaviour therapy is the best‐validated treatment for PTSD, it has stagnated over recent decades, and only two‐thirds of PTSD patients respond adequately to this intervention. Moreover, most people with PTSD do not access evidence‐based treatment, and this situation is much worse in low‐ and middle‐income countries. Identifying processes that can overcome these major barriers to better management of people with PTSD remains an outstanding challenge.

JOG:
Observer-Rated and Self-Report methods of measuring patient PTSD/anxiety levels were used to assess patient PTSD levels across three groups of fifty patients. Each group has an equal numbers of Male and Female participant, twenty-five each. 

The age range for this hypothetical situation doesn't matter as it hasn't been recorded, but I suppose if this was a real world situation there might be some ethical issues, but it's not sure it's all made up as far as I know. 

I suppose the real lesson in any college assignment is how much unnecessary information we can filter through, as a student, before you stop caring because life's too short.

There was a 12 week interval between the pre and post trial measurements, or the trial start and end. The types of groups really don't matter, they are just variable names for me. They might as well be x, y and z. What the study is about doesn't matter either, they'll make more readable labels for graphs later. All the waffle about PTSD kind of takes away from the fact this is basically just a numbers and graphs exercise, so I'm not wasting my time rehashing the stuff from the CA spec.

**Groups:**

* Control (No VR)
* Static (Non-Animated VR)
* Animated (Animated VR)

**Gender: (Data Limited)**

* Male
* Female

**Measurements:**

* Self-Report:
    + Start PTSD Level (pre_trial_cpss)
    + End PTSD Level (post_trial_cpss)
* Observer-Rated:
    + Start PTSD Level (pre_trial_or)
    + End PTSD Level (post_trial_or)


## Rationale
MG:
CBT addresses the here and now of PTSD symptoms, it doesn’t necessarily go back over the initial cause of the PTSD. That is something other types of therapies can explore with you. CBT helps you to adjust and re-align negative thoughts, feelings and behaviours that stop you from enjoying a good quality of life.
We aim to explore the effectiveness of virtual reality-assisted cognitive behavioural therapy (VRCBT) in the treatment of patients with PTSD. We further compare the therapeutic effect of VRCBT with that of standard cognitive behavioural therapy (CBT).

JOR:
Kind of simple. We only have a small amount of information, the change in PTSD Levels after 12 weeks, with one possible variable gender that may influence the outcome. Using 3 different treatment methods, with 2 different ratings. 

So, we would expect there is either some change, or no change after 12 weeks. Each of the methods may have different outcomes, and the gender of the patients may be a factor in how they respond to each treatment, or to treatment in general by any of the 3 treatments.

## Hypothesis
MG:
Null hypothesis: there is no difference in terms of CPSS scores betweeen the animated group and control group, or the static group and the control group
Alternative hypothesis: there is a significant difference in terms of CPSS scores between the animated group and control group, or the static group and the control group


JOR:
Average PTSD levels for both observer rated and self-reported should improve over a 12 week period.

H0 (Null hypothesis): 

H1 (Alternative): 

# Method

## Participants
The participants involved in the study were all young adults with PTSD in the age range of 18 - 25 years. There were 150 students, equally divided into 75 males and 75 females.

## Design
The patients were allocated into one of three groups; 
- Control, static (traditional CBT, no VR)
- Static (non-animated model content, VR) 
- Animated (animated model content, VR)

Non-animated means the objects in the virtual environment would be still and not interact.
Animated means when you add objects onto the Jungian Sandplay they would interact and move around the environment

## Materials
The 3 groups underwent 12 weeks of treatment for 50 minutes per week with a therapist. During that time the patients either underwent traditional CBT; or used one of the two (quality) versions of the VR app.

## Procedure
During the study, PTSD was assessed using observer-rated (i.e., therapist-rated) and self-report (i.e., Child PTSD Symptom Scale Self-Report Version (CPSS-SR)) measurements, respectively. Both measurements are scaled to the range of 0 to 10. Measurements were taken at the start and end of the study.

# Results

## Descriptive Statistics

plots/charts
e.g. boxplot

* clean it - get rid of errors
* identify outliers
* plot the graph
  - bar plots
  - box plots
  - histograms
  - scatterplots
  
* once we plot, look at things we can begin to explore data
* begin to develop hypothesis
* then run an experiment
* infer something from experiment


Measures of central tendency

* mean

##### mean()
##### trim - drops values - 0.3, drops 3 values from each end
##### <data>.mean <- mean(<variable>, trim = 0.3)


```{r}
pre_trial_cpss_mean <- mean(data5$pre_trial_cpss)
pre_trial_cpss_mean

pre_trial_cpss_std <- sd(data5$pre_trial_cpss)
pre_trial_cpss_std
pre_trial_cpss <- rnorm(150, pre_trial_cpss_mean, pre_trial_cpss_std)
hist(pre_trial_cpss)

# clean data? 1 value missing for this variable

post_trial_cpss_mean <- mean(data5$post_trial_cpss)
post_trial_cpss_mean
#post_trial_cpss_std <- sd(data5$post_trial_cpss)
#post_trial_cpss_std
#post_trial_cpss <- rnorm(150, post_trial_cpss_mean, post_trial_cpss_std)
#hist(post_trial_cpss)

pre_trial_or_mean <- mean(data5$pre_trial_or)
pre_trial_or_mean
pre_trial_or_sd <- sd(data5$pre_trial_or)
pre_trial_or_sd
pre_trial_or <- rnorm(150, pre_trial_or_mean, pre_trial_or_sd)
hist(pre_trial_or)

post_trial_or_mean <- mean(data5$post_trial_or)
post_trial_or_mean
post_trial_or_sd <- sd(data5$post_trial_or)
post_trial_or_sd
post_trial_or <- rnorm(150, post_trial_or_mean, post_trial_or_sd)
hist(post_trial_or)


```

* median

```{r}
pre_trial_cpss_mode <- mean(data5$pre_trial_cpss)
pre_trial_cpss_mean
pre_trial_cpss_std <- sd(data5$pre_trial_cpss)
pre_trial_cpss_std
pre_trial_cpss <- rnorm(150, pre_trial_cpss_mean, pre_trial_cpss_std)
hist(pre_trial_cpss)

# clean data? 1 value missing for this variable
#post_trial_cpss_mean <- mean(data5$post_trial_cpss)
# na.rm - removes null values
post_trial_cpss_mean <- mean(data5$post_trial_cpss, na.rm = TRUE)
#post_trial_cpss_mean
#post_trial_cpss_std <- sd(data5$post_trial_cpss)
#post_trial_cpss_std
#post_trial_cpss <- rnorm(150, post_trial_cpss_mean, post_trial_cpss_std)
#hist(post_trial_cpss)

pre_trial_or_mean <- mean(data5$pre_trial_or)
pre_trial_or_mean
pre_trial_or_sd <- sd(data5$pre_trial_or)
pre_trial_or_sd
pre_trial_or <- rnorm(150, pre_trial_or_mean, pre_trial_or_sd)
hist(pre_trial_or)

post_trial_or_mean <- mean(data5$post_trial_or)
post_trial_or_mean
post_trial_or_sd <- sd(data5$post_trial_or)
post_trial_or_sd
post_trial_or <- rnorm(150, post_trial_or_mean, post_trial_or_sd)
hist(post_trial_or)


```

* mode

Measures of variability

* standard deviation
* variance
* minimum and maximum variables

## Inferential Statistics

chapter 7?

Taking a sample and inferring something in the population
Probability of something
Generalise sample statistic to population
- mean response in population is x
- believe with % certainty

Histograms
Scatter Plots

## Statistical Tests
## Magnitude and Direction of Results

# Discussion

## Outline Findings and Relation to the Hypothesis

??? Hypothesis testing conclusion ???
change in anxiety over time +
compare 3 groups

## Limitations (If confounding variables are clearly identified by your group)

# References

**OpenIntro Statistics:**

M., D., D., C. and Çetinkaya-Rundel, M., 2019. OpenIntro Statistics. OpenIntro, Incorporated.

OpenIntro Statistics. 2022. OpenIntro Statistics. [ONLINE] Available at: https://www.openintro.org/book/os/. [Accessed 21 December 2022].

**Tutorialspoint:**

R Tutorial. 2022. R Tutorial. [ONLINE] Available at: https://www.tutorialspoint.com/r/index.htm. [Accessed 21 December 2022].

**Table of Contents:**

Yihui Xie, J. J. Allaire, Garrett Grolemund. 2022. 3.1 HTML document | R Markdown: The Definitive Guide. [ONLINE] Available at: https://bookdown.org/yihui/rmarkdown/html-document.html. [Accessed 21 December 2022].

**Cleaning Data**

R Programming 101 (YouTube). 2022. Clean your data with R. R programming for beginners. - YouTube. [ONLINE] Available at: https://youtu.be/sV5lwAJ7vnQ. [Accessed 22 December 2022].


**T Test**

R Programming 101 (YouTube). 2022. Doing a t-test using R programming (in 4 minutes) - YouTube. [ONLINE] Available at: https://www.youtube.com/watch?v=x1RFWHV2VUU. [Accessed 23 December 2022].


**Markdown**

Markdown Basics. 2022. Markdown Basics. [ONLINE] Available at: https://rmarkdown.rstudio.com/authoring_basics.html. [Accessed 23 December 2022].

**PTSD and CBT research**
https://www.ncbi.nlm.nih.gov/pmc/articles/PMC6732680/
https://www.ptsduk.org/treatment-help/cognitive-behavioural-therapy-cbt/

# Misc

## column names

* gender
* test_group
* pre_trial_cpss
* post_trial_cpss
* pre_trial_or
* post_trial_or


```{r}
unique_values_in_gender <- unique(data5$gender)
unique_values_in_gender
```

"Feale" needs to be changed to "Female"

```{r}
unique_values_in_test_group <- unique(data5$test_group)
unique_values_in_test_group

```

"Anmated" needs to be changed to "Animated"

## cleaning



## Random Plots

Testing plots

```{r data5, echo=FALSE}
plot(data5$pre_trial_cpss)
#plot(data5$post_trial_cpss) # clean data
plot(data5$pre_trial_or)
plot(data5$post_trial_or)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
