---
title: "DAIE Pair Project"
author: "Joe O'Regan and Michael Galbraith"
date: "2022-12-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Data Analytics and Immersive Environments - Pair Project
### Joe O'Regan and Michael Galbraith

**Repo:** <https://github.com/joeaoregan/2022_DAIE_GCA_JOR_MG>

## 2. Table of Contents

## 3. Abstract

### - Aim and Rationale
### - Participants and Setting
### - Experiment Design
### - Results Gathering
### - Major Findings
### - Findings / Implications

## 4. Introduction

### - Topic and Context
### - Rationale
### - Hypothesis


## 5. Method

### - Participants
### - Design
### - Materials
### - Procedure

## 6. Results

### - Descriptive Statistics

Measures of central tendency

* mean
* median
* mode

Measures of variability

* standard deviation
* variance
* minimum and maximum variables

### - Inferential Statistics
### - Statistical Tests
### - Magnitude and Direction of Results

## 7. Discussion

### - Outline Findings and Relation to the Hypothesis
### - Limitations (If confounding variables are clearly identified by your group)

## 8. References

M., D., D., C. and Çetinkaya-Rundel, M., 2019. OpenIntro Statistics. OpenIntro, Incorporated.

OpenIntro Statistics. 2022. OpenIntro Statistics. [ONLINE] Available at: https://www.openintro.org/book/os/. [Accessed 21 December 2022].



## Misc



```{r Load & View, echo=TRUE}

library(readr)
data5 <-read_csv("daie_ca3_data_5.csv")
View(data5)
#head(data5, 10)
```



```{r}
pre_trial_cpss_mean <- mean(data5$pre_trial_cpss)
pre_trial_cpss_mean
pre_trial_cpss_std <- sd(data5$pre_trial_cpss)
pre_trial_cpss_std
pre_trial_cpss <- rnorm(150, pre_trial_cpss_mean, pre_trial_cpss_std)
hist(pre_trial_cpss)

# clean data? 1 value missing for this variable
#post_trial_cpss_mean <- mean(data5$post_trial_cpss)
#post_trial_cpss_mean
#post_trial_cpss_std <- sd(data5$post_trial_cpss)
#post_trial_cpss_std
#post_trial_cpss <- rnorm(150, post_trial_cpss_mean, post_trial_cpss_std)
#hist(post_trial_cpss)

pre_trial_or_mean <- mean(data5$pre_trial_or)
pre_trial_or_mean
pre_trial_or_sd <- sd(data5$pre_trial_or)
pre_trial_or_sd
pre_trial_or <- rnorm(150, pre_trial_or_mean, pre_trial_or_sd)
hist(pre_trial_or)

post_trial_or_mean <- mean(data5$post_trial_or)
post_trial_or_mean
post_trial_or_sd <- sd(data5$post_trial_or)
post_trial_or_sd
post_trial_or <- rnorm(150, post_trial_or_mean, post_trial_or_sd)
hist(post_trial_or)


```


## Random Plots

Testing plots

```{r data5, echo=FALSE}
plot(data5$pre_trial_cpss)
#plot(data5$post_trial_cpss) # clean data
plot(data5$pre_trial_or)
plot(data5$post_trial_or)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
