---
title: "1. DAIE Pair Project"
author: "Joe O'Regan and Michael Galbraith"
date: "2022-12-12"
output:
  html_notebook:
    theme: darkly
    toc: yes
    toc_depth: 2
    #number_sections: true # numbers instead of bullet points
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r Load & View, echo=TRUE}

library(readr)
data5 <-read_csv("daie_ca3_data_5.csv")
View(data5)
#head(data5, 10)
```

**Repo:** <https://github.com/joeaoregan/2022_DAIE_GCA_JOR_MG>

# Title
# Table of Contents
# Abstract

## Aim and Rationale

## Participants and Setting

## Experiment Design

## Results Gathering

## Major Findings

## Findings / Implications


# Introduction

## Topic and Context
## Rationale
## Hypothesis


# Method

## Participants
## Design
## Materials
## Procedure

# Results

## Descriptive Statistics

Measures of central tendency

* mean

##### mean()
##### trim - drops values - 0.3, drops 3 values from each end
##### <data>.mean <- mean(<variable>, trim = 0.3)


```{r}
pre_trial_cpss_mean <- mean(data5$pre_trial_cpss)
pre_trial_cpss_mean
pre_trial_cpss_std <- sd(data5$pre_trial_cpss)
pre_trial_cpss_std
pre_trial_cpss <- rnorm(150, pre_trial_cpss_mean, pre_trial_cpss_std)
hist(pre_trial_cpss)

# clean data? 1 value missing for this variable
#post_trial_cpss_mean <- mean(data5$post_trial_cpss)
#post_trial_cpss_mean
#post_trial_cpss_std <- sd(data5$post_trial_cpss)
#post_trial_cpss_std
#post_trial_cpss <- rnorm(150, post_trial_cpss_mean, post_trial_cpss_std)
#hist(post_trial_cpss)

pre_trial_or_mean <- mean(data5$pre_trial_or)
pre_trial_or_mean
pre_trial_or_sd <- sd(data5$pre_trial_or)
pre_trial_or_sd
pre_trial_or <- rnorm(150, pre_trial_or_mean, pre_trial_or_sd)
hist(pre_trial_or)

post_trial_or_mean <- mean(data5$post_trial_or)
post_trial_or_mean
post_trial_or_sd <- sd(data5$post_trial_or)
post_trial_or_sd
post_trial_or <- rnorm(150, post_trial_or_mean, post_trial_or_sd)
hist(post_trial_or)


```

* median

```{r}
pre_trial_cpss_mode <- mean(data5$pre_trial_cpss)
pre_trial_cpss_mean
pre_trial_cpss_std <- sd(data5$pre_trial_cpss)
pre_trial_cpss_std
pre_trial_cpss <- rnorm(150, pre_trial_cpss_mean, pre_trial_cpss_std)
hist(pre_trial_cpss)

# clean data? 1 value missing for this variable
#post_trial_cpss_mean <- mean(data5$post_trial_cpss)
# na.rm - removes null values
post_trial_cpss_mean <- mean(data5$post_trial_cpss, na.rm = TRUE)
#post_trial_cpss_mean
#post_trial_cpss_std <- sd(data5$post_trial_cpss)
#post_trial_cpss_std
#post_trial_cpss <- rnorm(150, post_trial_cpss_mean, post_trial_cpss_std)
#hist(post_trial_cpss)

pre_trial_or_mean <- mean(data5$pre_trial_or)
pre_trial_or_mean
pre_trial_or_sd <- sd(data5$pre_trial_or)
pre_trial_or_sd
pre_trial_or <- rnorm(150, pre_trial_or_mean, pre_trial_or_sd)
hist(pre_trial_or)

post_trial_or_mean <- mean(data5$post_trial_or)
post_trial_or_mean
post_trial_or_sd <- sd(data5$post_trial_or)
post_trial_or_sd
post_trial_or <- rnorm(150, post_trial_or_mean, post_trial_or_sd)
hist(post_trial_or)


```

* mode

Measures of variability

* standard deviation
* variance
* minimum and maximum variables

## Inferential Statistics

chapter 7?

## Statistical Tests
## Magnitude and Direction of Results

# Discussion

## Outline Findings and Relation to the Hypothesis
## Limitations (If confounding variables are clearly identified by your group)

# References

**OpenIntro Statistics:**

M., D., D., C. and Çetinkaya-Rundel, M., 2019. OpenIntro Statistics. OpenIntro, Incorporated.

OpenIntro Statistics. 2022. OpenIntro Statistics. [ONLINE] Available at: https://www.openintro.org/book/os/. [Accessed 21 December 2022].

**Tutorialspoint:**

R Tutorial. 2022. R Tutorial. [ONLINE] Available at: https://www.tutorialspoint.com/r/index.htm. [Accessed 21 December 2022].

**Table of Contents:**

Yihui Xie, J. J. Allaire, Garrett Grolemund. 2022. 3.1 HTML document | R Markdown: The Definitive Guide. [ONLINE] Available at: https://bookdown.org/yihui/rmarkdown/html-document.html. [Accessed 21 December 2022].

# Misc

## Random Plots

Testing plots

```{r data5, echo=FALSE}
plot(data5$pre_trial_cpss)
#plot(data5$post_trial_cpss) # clean data
plot(data5$pre_trial_or)
plot(data5$post_trial_or)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
